---
title: "Initial Korean covid 19"
author: "Wayne Jarvis"
date: "Fall 2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Graphical results - use ggplot commands to do this  

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

I observed that the number of people alive are higher than the number of people dead. I found that in the chart of dead people 2/3 of it were men and 1/3 were women. This tells us that the disease killed men twice as much as women.

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
prop.table(table1)
prop.table(table1)
```

The percentage of women dying from COVID is 1.4% while for men it's 3%. This percentage is evidence that the disease is twice as fatal towards men. It also matches with the data observed from our Graphical Results.

## Inferential results 

```{r}
chisq.test(table1)
fisher.test(table1)
```

I found that the odds of men dying from COVID in South Korea was 2.3 times greater than women. This matches with past results as we saw that most dead from COVID were men. The P-value (0.00009) is evidence that men have a higher chance of dying than women.

## Conclusions

This assignment allowed me to analyze data from South Korea. It was nice to observe real world data instead of hypothetical data. It was difficult to swallow that these were real people, especially since twice as many men died. In particular, I analyzed data regarding COVID deaths in South Korea, and how sex played a role in death distribution. There are several pieces of evidence showing how men died twice as much. The first is taken from Graphical Results, which shows that 2/3 of the people dead were men while 1/3 were women. The Numerical Results show us that 1.4% of women died while 3% of men died. Lastly, in Inferential Results, I noticed that men died 2.3 times as much as women. In conclusion, the death toll was immense, affecting men twice as much as women; a tragedy.
